---
title: "Lab 04 - La Quinta is Spanish for next to Denny's, Pt. 1"
author: "Rowan Kemmerly"
date: "2.7.23"
output: github_document
---

### Load packages and data

```{r}
# install.packages("devtools")
devtools::install_github("rstudio-education/dsbox")
```

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox) 
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
```

### Exercise 1

```{r dataset-info}
glimpse(dennys)
```

The Denny's dataset has 1643 rows and 6 columns: each row represents a Denny's location, and each column represents the 6 variables of location street address, city, state, zip code, longitude, and latitude. 

### Exercise 2

```{r dataset-info}
glimpse(laquinta)
```

The La Quinta dataset has 909 rows and 6 columns: each row represents a La Quinta location, and each column represents the 6 variables of location street address, city, state, zip code, longitude, and latitude. 


### Exercise 3

There are 17 La Quinta locations outside of the United States in Canada, Mexico, China, New Zealand, Turkey, United Arab Emirates, Chile, and Colombia. There are no Denny's locations outside of the US. 

### Exercise 4

```{r view-data}
view(laquinta)
view(dennys)
```

To find out whether La Quinta or Denny's have locations outside of the US using the dataset, you could filter for states that are not one of the abbreviations of 50 (51) US states. That could probably be done using the help of our "states" dataset!

### Exercise 5

Seeing whether there are Denny's locations outside of the US with code:

```{r dennys-filter}
dennys %>%
  filter(!(state %in% states$abbreviation))
```

Confirmed that there are no Denny's locations outside of the US!

### Exercise 6

Adding a "country" variable to the Denny's dataset:

```{r dennys-country}
dennys %>%
  mutate(country = "United States")
```

### Exercise 7
 
Seeing whether there are La Quinta locations outside of the US with code:

```{r la-quinta-filter}
laquinta %>%
  filter(!(state %in% states$abbreviation))
```

So it looks like this dataset only contains 14 of La Quinta's international locations. The ones in this dataset are located in Mexico (10 locations, 1 of which is now permanently closed, and 3 of which are now CHN hotels instead), Colombia, Honduras, and Canada (2 locations):

Carretera Panamericana Sur KM 12 (Mexico)
Av. Tulum Mza. 14 S.M. 4 Lote 2	(Mexico, permanently closed)
Ejercito Nacional 8211 (Mexico)
Blvd. Aeropuerto 4001	(Mexico)
Carrera 38 # 26-13 Avenida las Palmas con Loma de San Julian El Poblado	(Colombia)
AV. PINO SUAREZ No. 1001 (Mexico)
Av. Fidel Velazquez #3000 Col. Central (Mexico)
63 King Street East	(Canada)
Calle Las Torres-1 Colonia Reforma (Mexico)
Blvd. Audi N. 3 Ciudad Modelo (Mexico)
Ave. Zeta del Cochero No 407 (Mexico)
Av. Benito Juarez 1230 B (Carretera 57) Col. Valle Dorado Zona Hotelera	(Mexico)
Blvd. Fuerza Armadas (Honduras)
8640 Alexandra Rd (Canada)

### Exercise 8

Now I'm adding a "country" variable based on state abbreviations for each case.

```{r adding-country-variable}
laquinta <- laquinta %>%
  mutate(country = case_when(
    state %in% state.abb     ~ "United States",
    state %in% c("ON", "BC") ~ "Canada",
    state == "ANT"           ~ "Colombia",
    state == "FM"            ~ "Honduras",
    state %in% c("AG","QR", "CH", "NL", "VE","PU","SL") ~ "Mexico"
  ))
```

Now I'm filtering the La Quinta dataset for just locations that are in the US.

```{r la-quinta-just-US}
laquinta <- laquinta %>%
  filter(country == "United States")
```

### Exercise 9

Let's now look at how many Denny's and La Quintas are in each state!

```{r dn-frequencies}
dn_freq <- dennys %>%
  count(state) %>%
  inner_join(states, by = c("state" = "abbreviation")) %>%
  arrange(desc(n))

```

```{r lq-frequencies}

lq_freq <- laquinta %>%
  count(state) %>%
  inner_join(states, by = c("state" = "abbreviation")) %>%
  arrange(name)
```

So it looks like California has the most Denny's locations (makes sense, since it's such a large state), and Delaware has the least (which also makes sense, since it's such a small state). For La Quinta locations, Texas has by far the most (not entirely sure why that's the case) and Delaware, Hawaii, and DC all have the least (since they all have 0 locations and aren't in this data frame — Maine has 1). 


### Exercise 9

First I'm creating a new data frame that has both the Denny's and La Quinta data. 

```{r creating-combined-dataset}

dn <- dennys %>%
  mutate(establishment = "Denny's")
lq <- laquinta %>%
  mutate(establishment = "La Quinta")

dn_lq <- bind_rows(dn, lq)

```

```{r plot-1}

ggplot(dn_lq, aes(x = longitude,
                  y = latitude,
                  color = establishment)) +
  geom_point()

```

Wow, it's the shape of the US! 

I'm going to compute a new variable to determine which states have the most Denny's and La Quinta locations per thousand square miles.

```{r per-1000-miles}

dn_freq %>%
  mutate(per1000sqmi = (n/area)*1000) %>%
  arrange(desc(per1000sqmi))

lq_freq %>%
  mutate(per1000sqmi = (n/area)*1000) %>%
  arrange(desc(per1000sqmi))


```

DC has the most Denny's per 1000 square miles, and Rhode Island has the most La Quintas per square 1000 miles!




